Exploring the Traveling Salesman Problem

### 2017-05-02 paper/paper.tex
finish algorithm comparison

% Packages
% ---
\usepackage[
  left=3cm,
  right=3cm,
  top=2cm,
  bottom=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{clrscode3e}
\usepackage{blkarray}
\usepackage{booktabs}
\usepackage{parskip}
\usepackage{mathptmx}
\usepackage{bigstrut}
\usepackage{mathptmx}
\usepackage{minted}
\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
\usepackage{apacite}
%\usepackage{natbib}

@@

\subsection{Christofides}
Christofide's algorithm is an $O(n^3)$ approximation algorithm that guarentees
the cost of its solution will be strictly within $3/2$ of the optimal solution
cost.  The algorithm only works on a symmetric distance matrix that satisfies
the triangle inequality \cite{nicos}.  That is, a distance matrix $D=d_{ij}$
satisfies the triangle inequality if the shortest path to reach every $j$ from
$i$ involves traveling directly to $j$ from $i$, rather than through a path that
includes more than two verticies.
\par

The algorithm works as follows: Given a distance matrix $D=d_{ij}$,

>>>>

% Packages
% ---
\usepackage[
  left=3cm,
  right=3cm,
  top=2cm,
  bottom=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{clrscode3e}
\usepackage{blkarray}
\usepackage{booktabs}
\usepackage{parskip}
\usepackage{mathptmx}
\usepackage{bigstrut}
\usepackage{mathptmx}
\usepackage{minted}
\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
\usepackage{apacite}
\usepackage{tikz}

@@

\subsection{Christofides}
Christofide's algorithm is an $O(n^3)$ approximation algorithm that guarentees
the cost of its solution will be strictly within $3/2$ of the optimal solution
cost.  The algorithm only works on a symmetric distance matrix that satisfies
the triangle inequality \cite{nicos}.  That is, a distance matrix $D=d_{ij}$
may be used if the shortest path to reach every $j$ from
$i$ involves traveling directly to $j$ from $i$, rather than through a path that
includes more than two verticies.
\par

The algorithm works as follows: Given a distance matrix $D=d_{ij}$ and size $n$,
let $G=(V,E)$ where $v_i=i$ for all $i\in [1,\dots,n]$ and $e_{ij}=d_{ij}$.  The
optimal tour starting at $v_1$ is obtained using the following steps:
\begin{enumerate}
  \item Generate a minimum spanning tree, $MST$, rooted at $v_1$ using Prim's algorithm.
  \item Perform a preorder traversal on $MST$ and append the verticies in the order they 
    are visited to list a $T$.
  \item For each verticy in $T$, append it to $T'$ only when it is first visited.  This 
    removes duplicates from $T$ resulting in the optimal tour.
\end{enumerate}

The last step holds since $G$ satisfies the triangle inequality.

\subsection{Nearest Neighbor}
The Nearest Neighbor algorithm is a greedy approximation algorithm
for solving the TSP\@.  It starts at an arbitrary city, then visits
the closest city until each city has been visited once.  After each
visit, the city is added to the optimal tour $T$.  Once each node
has been visited, the starting city is appended to $T$ and $T$ is
returned. \cite{bellmore}
\par

The steps for this algorithm are as follows:
\begin{enumerate}
  \item Pick a starting city $s$, let it be the current city $x$.
  \item Add $x$ to the optimal tour $T$ and mark $x$ as visited.
  \item Determine the closest unvisited city $y$ from $x$.  If no
    unvisited cities exist, goto step 5.
  \item assign $x$ to $y$, repeat step 2.
  \item Add $s$ to $T$, return $T$.
\end{enumerate}

\subsection{Comparing the Three}
We have examined one exact algorithm -- Held-Karp -- and two approximation
algorithms --Christofides and Nearest Neighbor.  The approximation algorithms
have the advantage of being significantly faster than the Held-Karp algorithm,
but they aren't applicable in every case, and they don't always return the optimal
tour for cases in which they are applicable.  The Held-Karp algorithm, on the other
hand, will return the optimal tour in every TSP instance, provided the number of
cities is small enough to allow the algorithm to finish execution in a reasonable
time.
\par

Christofide's algorithm and the Nearest Neighbor algorithm both perform at an
acceptable level (that is, they return the optimal tour enough of the time) for
Euclidean TSP instances.  If a TSP instance is Euclidean, then it satisfies the
triangle inequality.  This property is necessary for Christofide's algorithm to
return the correct tour \cite{nicos} and it greatly increases the probability
that the Nearest Neighbor algorithm will as well \cite{gutin}.
\par

For asymmetric and general symmetric TSP instances, Christofide's algorithm will
never return the optimal tour.  The Nearest Neighbor algorithm, on the other hand,
can find an optimal tour, but it's highly unlikely.  Additionally, for every $n\gte 2$
there exists at least one asymmetric or general symmetric TSP instance for which the
Nearest Neighbor algorithm returns the \textit{worst} tour. \cite{gutin}.  We are
optimizing for correctness over speed and problem size in this paper and will continue
to explore the Held--Karp algorithm.

\subsection{Held-Karp Pseudocode}
